---
title: "R Assignment 1"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  markdown:
    wrap: 72
---

The Behavioral Risk Factor Surveillance System (BRFSS) is an annual
telephone survey of 350,000 people in the United States. As its name
implies, the BRFSS is designed to identify risk factors in the adult
population and report emerging health trends. For example, respondents
are asked about their diet and weekly physical activity, their HIV/AIDS
status, possible tobacco use, and even their level of healthcare
coverage.

The BRFSS Web site (<http://www.cdc.gov/brfss>) contains a complete
description of the survey, including the research questions that
motivate the study and other interesting results. We will focus on a
random sample of people from the BRFSS survey conducted in 2000. While
there are over 200 variables in this data set, we will work with a small
subset.

## Loading the **dataset**

Begin by loading the dataset into your R workspace by running the
command in the chunk below:

```{r}
source("http://www.openintro.org/stat/data/cdc.R")
```

The file cdc.R is an R-**data set**, as such it has a special format,
and we can use certain commands to access the data.

> **Note**: for other types of formats, we may have to use other
> commands

## Viewing the data set

Would it be a good idea to *view* this entire data set? The command
`View` will create a table with all entries of the data set. See what
happens when you run the chunk below ( the output may open in a separate
tab. Make sure to come back to this tab to continue.)

```{r}
View(cdc)
```

Ok, perhaps it's a better idea to read parts of the data set. For
example, if we want to see the variables in the set, we can run the
command `names`. Do it in the chink below to find out what variables are
stored in this data set.

```{r}
names(cdc)
```

We can also view, for example, the first 5 rows of the data set. Run the
command below to see the output

```{r}
head(cdc, 5)
```

Similarly, we can also view the last 5 rows by writing the command
`tail(cdc, 5)`

# Viewing parts of the data set

What if we want to see only some variables ( columns) but not others?
The command below returns the columns 2-thorugh-5 of the data set only

```{r}
cdc[2:5]
```

## Assignation

Sometimes it will be useful to make a different data frame, and save it
under a different name. In R, this is called an **assigning an object to
a variable**, or an *assignment*. We do this by typing

`new_object -< old_object`

For example, the following command will create a new data set, called
mycdc, that includes only the first 1000 rows if the original data set

```{r}
mycdc <- head(cdc, 1000)
```

## Ordering data by a variable

If we want to reorder the data according to the values of one of the
columns, we can do that by using the command `order( , )` the syntax is

`data_frame[order(data_frame&variable), ]`

Run the chinks of code below to see it in action

```{r}
ordered <- cdc[order(mycdc$height), ]
```

Let's now see the first 10 rows of this new data set. Run the chunk of
code below.

```{r}
head(ordered, 10)
```

# Questions

## Q1

List all variables in the data set cdc.R. For each variable, determine
if it is Categorical, or numerical, and if numerical, if it is discrete
or continuous.

```{r}
names(cdc)
```

-   Gen health: **Categorical**

-   Exerany: **categorical**

-   Hlthplan; **Categorical**

-   Smoke: **Categorical**

-   Height: **Numerical: continuous**

-   Weight" **Numerical: Continuous**

-   WtDesired: **Numerical Continous**

-   Age: **Numerical Discrete**

-   Gender: **Categorical**

## Q3

How many people were used in this survey?

20,000

## Q2

Create a new data set that has only the last 5 columns, and only the
first 5000 rows of the data set cdc.R

```{r}
lastcdc<- cdc[5:9]
head(lastcdc, 5000)
```

## Q4

Order the data set you have in Q3 by age. Save it as a new data set.

```{r}
ordy <- lastcdc[order(lastcdc$age), ]
head(ordy, 5000)
```
